Stochastic blockmodeling of relational event dynamics

### 2012-05-31 docs/intro.tex
Clean up second paragraph.

Statistical methods for analyzing network data are becoming increasingly useful for studying phenomena ranging from online social behavior to protein  interactions \cite{Goldenberg2009}.
In such settings we often observe events occurring between nodes over time, and a common goal is to understand the differences in the total number of interactions --- perhaps with respect to known information about the nodes.
%In this case, we have a count for the total activity of each dyad, and one may analyze it has a static network. 
%Methods for static networks such as a 
Stochastic blockmodels \cite{Nowicki2001, Kemp, Ishiguro2010, Rodriguez} are a class of techniques which can model this heterogeneity by assuming latent subgroups of nodes have similar tendencies in their group-wise rate of interaction.

Network-based event data are powerful, however, because they enable the study of the processes governing these events.
Theories about the dynamics among nodes will often be difficult to study when only using the marginal rate of events, as illustrated in Figure \ref{fig:introexample}.  
For example, a common aspect of human communication is \emph{reciprocity}: person A speaks to person B, then B has a higher propensity to speak to A.  
Through an event-based analysis we can study the similarity --- and differences --- in the dynamics of nodes in the network. 
When studying interactions among students, for example, one might be interested in conversational tendencies are associated with covariates about the student or classroom, or based on unobserved groupings.  
In this context, a standard blockmodel only provides a means for studying the structure of marginal rates of communication, but not the shared tendencies in the dynamics of the sequence of events.

>>>>

Statistical methods for analyzing network data are becoming increasingly useful for studying phenomena ranging from online social behavior to protein  interactions \cite{Goldenberg2009}.
In such settings we often observe events occurring between nodes over time, and a common goal is to understand systematic differences in how often various interactions occur --- perhaps with respect to known information about the nodes.
Stochastic blockmodels \cite{Nowicki2001, Kemp, Ishiguro2010, Rodriguez} are a class of techniques which can model this heterogeneity by assuming latent subgroups of nodes have similar tendencies in their group-wise rate of interaction.

Network-based event data are powerful, however, because they enable the study of the processes governing these events.
Theories about the dynamics among nodes will often be difficult to study when only using the marginal rate of events, as illustrated in Figure \ref{fig:introexample}.  
For example, a common aspect of human communication is \emph{reciprocity}: person A speaks to person B, then B has a higher propensity to speak to A.  
Through an event-based analysis we can study how nodes vary in their tendencies to behave in this manner.  %the similarity --- and differences --- in the dynamics of nodes in the network. 
This is particularly relevant when studying interactions among students in a classroom, where differences in conversational tendencies might be associated with covariates about the student or classroom, or based on unobserved groupings, and may have ramifications for education.
In such contexts, a standard blockmodel only provides a means for studying the structure of marginal rates of communication, but not the shared tendencies in the dynamics of the sequence of events.
